---
title: "Web Scraping with R and rvest"
author: "Aarushi Sahejpal"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Introduction to Web Scraping

- read HTML pages into R
- extract specific elements from a webpage
- extract tables from websites
- handle common web scraping challenges

## Load Required Libraries
```{r libraries}
library(rvest)      # For web scraping
library(dplyr)      # For data manipulation
library(tidyverse)  # For data wrangling and visualization
library(httr)       # For handling HTTP requests
```

---

# Part 1: Understanding HTML

HTML (HyperText Markup Language) is the standard language for creating web pages. When we scrape, we're extracting information from HTML documents.

**Key HTML concepts:**
- **Tags**: Elements like `<p>`, `<table>`, `<div>` that structure content
- **Attributes**: Properties of tags like `class` or `id`
- **Nodes**: Individual elements in the HTML tree

Learn more about [HTML basics](https://developer.mozilla.org/en-US/docs/Web/HTML)

---

# Part 2: Reading a Webpage

Let's start with a simple example by reading the rvest documentation page.
```{r simple-example}
# Read a simple webpage
html <- read_html("http://rvest.tidyverse.org/")

# Check what we got
class(html)
```

The `read_html()` function downloads the webpage and parses it into an R object we can work with.

---

# Part 3: Real-World Example - Bridge Data

Now let's work with real government data! We'll scrape the [Bridge Condition by Highway System 2024](https://www.fhwa.dot.gov/bridge/nbi/no10/condition24.cfm) page from FHWA.

## Step 1: Connect to the Website

Sometimes websites require special configurations to allow scraping. Government sites often have SSL certificate issues.
```{r connect-bridge}
# Define the URL
url <- "https://www.fhwa.dot.gov/bridge/nbi/no10/condition24.cfm"

# Create a session with SSL verification disabled
# (Note: only disable SSL when necessary for scraping)
session <- session(url, 
                   httr::config(ssl_verifypeer = FALSE))

# Read the webpage
webpage <- read_html(session)

# Verify it worked
class(webpage)
```

**Why use `session()`?**
- Handles cookies and authentication
- Maintains connection state
- More robust than simple `read_html()`

---

# Part 4: Exploring HTML Elements

Before extracting data, let's explore what's on the page.
```{r explore-elements}
# Extract all paragraph elements
paragraphs <- html_nodes(webpage, "p")

# Show the first few
head(paragraphs, 3)

# Extract all tables (without reading them yet)
tables <- html_nodes(webpage, "table")
length(tables)  # How many tables are on the page?
```

**Common CSS selectors:**
- `"p"` - all paragraph tags
- `"table"` - all tables
- `".classname"` - elements with a specific class
- `"#id"` - element with a specific id

---

# Part 5: Extracting Table Data

The bridge data is stored in a table. Let's extract it!
```{r extract-table}
# Extract all tables and get the first one
bridge_all <- html_table(html_nodes(webpage, "table"))[[1]]

# Look at the structure
head(bridge_all, 3)

# The first row often contains summary data or headers
# Let's remove it
bridges <- bridge_all[-1, ]

# View the cleaned data
head(bridges)
```

---

# Part 6: Data Inspection and Summary

Now that we have the data, let's explore it.
```{r inspect-data}
# Get column names
names(bridges)

# Check the structure
str(bridges)

# Summary statistics
summary(bridges)

# View dimensions
dim(bridges)
```


# Scrape Multiple Years (2022-2025)
```{r scrape-multiple-years}
# Function to scrape one year
scrape_bridge_year <- function(year) {
  url <- paste0("https://www.fhwa.dot.gov/bridge/nbi/no10/condition", year, ".cfm")
  message("Scraping year 20", year, "...")
  
  tryCatch({
    session <- session(url, httr::config(ssl_verifypeer = FALSE))
    webpage <- read_html(session)
    bridge_all <- html_table(html_nodes(webpage, "table"))[[1]]
    bridges <- bridge_all[-1, ]
    bridges$Year <- paste0("20", year)
    Sys.sleep(2)  # Be polite to the server
    return(bridges)
  }, error = function(e) {
    message("Failed to scrape year 20", year, ": ", e$message)
    return(NULL)
  })
}

# Scrape years 2022-2025
years <- c(22, 23, 24, 25)
bridge_list <- lapply(years, scrape_bridge_year)

# Remove any failed scrapes and combine
bridge_list <- bridge_list[!sapply(bridge_list, is.null)]
bridges_all_years <- bind_rows(bridge_list)

# Check results
print(paste("Total rows:", nrow(bridges_all_years)))
table(bridges_all_years$Year)

# Save combined data
write.csv(bridges_all_years, "bridges_2022_2025.csv", row.names = FALSE)
```

---

# Part 7: Data Cleaning

Scraped data often needs cleaning. Numeric columns are usually imported as characters.
```{r clean-data}
# Check for duplicate column names
names(bridges)

# See which ones are duplicated
names(bridges)[duplicated(names(bridges))]

# Fix duplicate names by making them unique
names(bridges) <- make.names(names(bridges), unique = TRUE)

# Now check again
names(bridges)

# Now you can clean the data
bridges_clean <- bridges %>%
  # Keep the first column as character, convert rest to numeric
  mutate(across(-1, ~as.numeric(gsub(",", "", .))))

# Check the cleaned structure
str(bridges_clean)
```


